---
title: "Palmer Penguins"
author: "Carlos Vasquez"
date: "09/12/2021"
output: 
  html_document:
    theme: cerulean
---

### Set up my environment

install.packages("palmerpenguins") #first time only

```{r message = FALSE, warning = FALSE, error = FALSE}
library(palmerpenguins)#data set pre-loaded
library("tidyverse")#package for data analysis
```

### view the data 

Let's check if the data was installed properly

```{r}
glimpse(penguins)
head(penguins)
View(penguins)#pulls up the table to view
```
### Pipe operator (ctrl+shift+m)
```{r}
penguins %>% 
  arrange(bill_length_mm)

penguins %>% 
  arrange(-bill_length_mm)

penguins %>% 
  group_by(island) %>% 
  drop_na() %>% 
  summarize(max_bill_length_mm = max(bill_length_mm))

penguins %>% 
  group_by(species, island) %>% 
  drop_na() %>% 
  summarize(max_bill_length_mm = max(bill_length_mm))
```
### Using ggplot2 and creating different charts
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g))

```

```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= bill_length_mm, y= bill_depth_mm))
```

#This scatter plot has data points that shows the the different species by color and shape
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g, color= species, shape= species))

```

#Alpha is helpful to make data points transparent for each species  
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g, alpha= species))

```

#Having the color purple outside of the aes function allows for all data points to reflect a single color
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g), color= "purple")
```

### facet function

#facet fuction is very helpful to display the different graphs for each data point
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g, color= species, shape= species))+
  facet_wrap(~species) 
```

#This facet function shows the different sex and species for the penguins (too busy and becomes harder to make an analysis from so many charts)
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g, color= species))+
  facet_grid(sex~species)
```

### geom_smooth()
#This creates a smooth line graph
```{r}
ggplot(data= penguins) + geom_smooth(mapping= aes(x= flipper_length_mm, y= body_mass_g))
```

### labs()

#Lets add some labels and annotations to some of these graphs
```{r}
ggplot(data= penguins) + geom_point(mapping= aes(x= flipper_length_mm, y= body_mass_g, color= species)) + geom_smooth(mapping= aes(x= flipper_length_mm, y= body_mass_g)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle = "Sample of Three Penguin Species", caption="Data Collected By Dr. Kristen Gorman")+
  annotate("text", x=220, y=3500,label="The Gentoos are the largest", color="purple", fontface= "bold", size= 4.5, angle=25)

```

#Let's save this graph as a .png 

ggsave('Three Penguins Species.png')#collapse environmet pane for complete graph
